---
title: "OM decay with and without exoenzymes"
author: "Drew"
format: gfm
editor: source
---

To make the point that exoenzymes doesn't just affect labile organic matter, I have made a simple, two-component model, which contains labile and recalcitrant organic matter. Both labile and recalcitrant OM decay, but in order to capture the effect of abiotic reactions, labile OM is slowly converted to recalcitrant OM ([Tranvik and Bertilsson 2001](https://onlinelibrary.wiley.com/doi/full/10.1046/j.1461-0248.2001.00245.x?casa_token=bPAvbpSsrowAAAAA%3Alg4u2CUXHlJGM0zy17ufH7Q-dG32LKE_l8B7H7Ed1IudfPYP0P1_KofEs2QkbdOdJ1A6LFC72Mooe5RG)). 

The model is thus:

![](https://latex.codecogs.com/svg.image?\frac{dG_l}{dt}&space;=&space;k_l&space;G_l&space;-&space;k_{l\rightarrow&space;r}G_l)

![](https://latex.codecogs.com/svg.image?\frac{dG_r}{dt}&space;=&space;k_rG_r&space;&plus;&space;k_{l\rightarrow&space;r}G_l)

I've solved these differential equations using the `deSolve` package for R, and the following rate constants:


\frac{dG_r}{dt} = k_rG_r + k_{l\rightarrow r}G_l


 | $k_l$ | $k_{l \rightarrow r}$ | $k_r$
--|--|--|--
enzymes | -0.01 | -0.002 | 0.005
no enzymes | -0.05 | -0.002 | 0.005

I'll spare you the code, but the model plot looks like this:


```{r, echo=FALSE, message = FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(deSolve)

# working off https://kinglab.eeb.lsa.umich.edu/480/nls/de.html
source("R/multi_G_model.R")

# Set the parameters for each case
k_l_r <- 0.02
k_r <- -0.005
params.enz <- c("k_l" = -0.1, "k_r" = k_r, "k_l_r" = k_l_r)
params.no.enz <- c("k_l" = -0.05, "k_r" = k_r, "k_l_r" = k_l_r)
times <- seq(from = 0, to = 100, by = 0.1)
xstart <- c(labile = 0.75, recalcitrant = 0.25)

d <- sim_G(params.enz, params.no.enz) 

theme <- theme_bw() + 
  theme(panel.grid = element_blank(),
        legend.position = "bottom")
p <- plot_ODE_data(d, theme=theme)
```

```{r, echo=FALSE, dpi=300, fig.height=2.5, fig.width=3.5}
print(p)
```

